---
title: "viz_and_eda"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggridges)
```

## create the weather data 

```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## create a ggplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

alternate way of amking this plot 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

saving initial plots: this is not used most of the time 

```{r}
scatterplot =     #creating a plot to save called scatterplot
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

scatterplot       #show plot
```


adding color...

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4)
```

why does 'aes' position matter? 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4) + 
  geom_smooth(se = FALSE)       #se = false takes away the error around line
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(alpha = .4)) + 
  geom_smooth(se = FALSE)
```

color applied in ggplot aes= will apply the color to everything that follows where as if it is applie in a later step it is only applied o that part (like geom_point it will only apply to the points not the line in geom smooth)


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(alpha = .4)) + 
  geom_smooth(se = FALSE) +
  facet_grid(~name) #this creates a plot for each category of the variable "name"
```





